---
title: "Preface/Chapter 1"
author: "ZS"
date: "May 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Welcome
## 1. Introduction

Programming: Import > Wrangle (Tidy & Transform) > Visualize, Model > Communicate

### 1.3 What you won't learn

* Big Data
* Python, Julia, and Friends
* Non-rectangular Data such as images, sounds, trees, and text
* Hypothesis Confirmation (book focuses on hypothesis generation)

### 1.4 Prerequisites

__Packages__: tidyverse

```{r echo=FALSE}
install.packages("tidyverse"); library(tidyverse); tidyverse_update()
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

### 1.6 Getting Help

__Stack Overflow__: Reprex (Reproduceable example)
Things to include in your reprex:
1. Packages
2. Data; using dput() to generate R code to recreate the dataframe:
  + Run dput(mtcars) in R
  + Copy the output
  + In my reproducible script, type mtcars <- then paste.
3. Code that's easy to read
  + using comments to clarify as needed
  + remove fluff that could distract from the problem

Try and find the smallest subset of your data that still reveals the problem.

