---
title: 'Part 1: Explore'
author: "ZS"
date: "May 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse"); 
tidyverse_update()
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

```{r Packages, echo=FALSE}
library(tidyverse)
library(nycflights13)
library(gapminder)
library(Lahman)
```

# Part 1: Explore
## 2. Introduction
1. Data Visualization
2. Data Transformation
3. Exploratory Data Analysis

## 3. Data Visualization
### 3.1 Introduction

Using ggplot2, which uses grammar of graphics for describing and building graphs
`r library(tidyverse)`

### 3.2 First Steps
Question: Do cars with big engines use more fuel than cars with small engines?
  + is the relationship positive/negative, linear/nonlinear?
  
```{r Plotting Basics, include=FALSE}
library(ggplot2)

mpg
str(mpg)

# To plot mpg
ggplot(data = mpg) + # this line tells R to use the mpg dataset
  geom_point(mapping = aes(x = displ, y = hwy)) # this line tells R to create a layer on the base plot, and what variable to plot on each axis

# So, really this is what you start with:
# ggplot(data = <DATA>) +
#   <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

# Plot hwy v cyl
ggplot(data = mpg) + # this line tells R to use the mpg dataset
  geom_point(mapping = aes(x = hwy, y = cyl))
# From this graph, you might say that cars with more cylinders have a lower highway mpg, though the relationship is not linear

# Plot class v drv
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
# This graph is not useful, because there does not seem to be a clear relationship between type of car (class) and what drive (front/rear/awd)
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = class))
# Plotting this way is a bit better, but still not super useful
```
### 3.3 Aesthetic mappings
To inform the graph of displ v hwy, we can add a third variable, like *class* to the two dimensional scatterplot by using an aesthetic. One way to think of this is if you were running a two-way ANOVA.

Aesthetics include things like size, shape, or color of your points. The same way you think of values of data, think of "levels" of aesthetic settings.

```{r Graph Aesthetics}
# Example: map the colors of your points to the "class" variable
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
# Here, you're mapping the name of the aesthetic to the variable inside aes()
# ggplot2 automatically assigns a color/level to each unique value of the variable => "scaling"

# We can do the same with scaling the class by size
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

# Mapping to alpha controls the transparency of the points
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Here, mapping with shape, which can deal with a max of 6 values
# If you're ever dealing with more than 6, consider using geom_point to map their levels manually
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

# Rather than setting aesthetic by class, could set for all points on graph manually
# Here the argument goes **outside** the aes() argument
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```
*Setting aesthetic manually*: To do so, you'll need to pick a value that makes sense per that aesthetic
+ name of a color as a character string --> "blue"
+ size of a point in millimeters (mm)
+ shape of a point as a number (0-20)

Exercises:
1) Which variables in mpg are categorical v quantitative?
`r ?mpg`
  + Categorical: manufacturer, model, trans, drv, fl, class, cyl
  + Quantitative: displ, year, cty, hwy
2) Map continuous variable to color, size, and shape 
```{r 3.3 Exercises}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cty, y = hwy, color = year, stroke = displ))

# Continuous variables cannot be mapped to shape, but color and size can!
# Stroke is the outline/fill of the shape. Size determines the size of the filled part, and Stroke determines the size of the stroke outline
```
### 3.4 Common Problems
When writing code that needs to continue to another line, denote with "+" at the *end* of the first line. 

Don't understand something about R's functioning? Try typing *?function_name* 

### 3.5 Facets
Rather than adding in additional variables into your plots, you can add *facets*, which are like subplots that each display one subset of the data

To facet your plot by a single variable, use facet_wrap(), where facet_wrap is followed by a formula "~ variable_name"
```{r Facets-One Variable}
library(ggplot2)
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2) # nrow denotes how many rows you want all facets to fit into, here it is two rows
```

You can also facet your plot on the combination of two variables using facet_grid.

```{r Facets-Two Variables}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl) # drv will be on the y axis, while cyl will be on the x-axis

# if you use "." on either side of the ~, you will facet columns or rows only, respectively ( row var ~ column var)
```

#### 3.5.1 Exercises
1. What happend if you facet a continuous variable (e.g. year)?
```{r 3.5.1 Exercises - # 1}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap (~ year, nrow = 4) # for year, it plots 1998 and 2008, which are the min and max years, respsectively, within the dataset

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap (~ cty, nrow = 5) # for cty, it graphs one graph per unique item in the continuous
# unique(mpg$cty) reveals 21 unique "bins" for city mpg, so it plots 21 plots to 5 rows
```

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r 3.5.1 Exercises - # 2}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)

# empty cells in plot indicate there is no data there. For instances, there are no 4-wheel or rear-wheel drive cars with 5 cylinders. There are also no cars with 7 cylinders
```
3. What plots does the following code make? What does "." do?
```{r 3.5.1 Exercises - # 3}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ .) # using "." plots just facets rows, not columns

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(. ~ cyl) # using "." plots just facets columns, not rows
```
4. Take the first plot in the section, what are the advantages/advantages to using faceting instead of the color aesthetic? How might you change things if you had a larger dataset?
```{r 5.3.1 Exercises - # 4}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2) # allows you to pick out trends within each class without other data crowding, would use this if I found that hwy was significantly different across displ and class

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) # better for seeing trends across classes, overall trends of data
```
5. nrow v ncol - can be used in facet wrap since you're telling it how many columns or rows to fit all levels of the wrapped variable within (e.g. 5 x 5). *However*, can't be used for facet grid, rows/columns of which is defined by the unique levels of the data for each variable involved.

6. 
